---
title: "Process [Insert Name] Dataset"
author: "Purdom Group"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(here)
# For relative file paths, this document expects `pop_data` as root folder
# `here::here()` should return "/accounts/projects/epurdom/singlecell/pop_data"

library(SeuratDisk) # For interface with h5Seurat format
source(here::here('Code','Process_Datasets','util_run_seurat_pipeline.R'))
source(here::here('Code','Process_Datasets','util_format_metadata.R'))
source(here::here('Code','Process_Datasets','util_get_cluster_proportions.R'))
```

# Set Up

```{r init_program,warning=FALSE,echo=FALSE}
# This notebook contains a template for processing a single-cell dataset into a Seurat object.

# Make sure to the `dataset_name` string assigned below leads to a valid `./data/Processed_Datasets/{dataset_name}` folder.
# The user must complete a processing program associated with this document which includes functions for creating a Seurat object from raw files and specification of processing choices.
# A template for the processing program can be found at `./Code/Process_Datasets/Programs/program_TEMPLATE.R`.
# The completed version should be saved at `./Code/Process_Datasets/Programs/program_{dataset_name}_{program_name}`.
# Bellow the Boolean `check_initial_seurat_save` determines whether to check for a "baseline" Seurat object containing the unmodified reads and metadata

dataset_name <- NULL # TODO: e.g. "ledergor"
program_name <- NULL #TODO: pick e.g. "eda"
primary_assay <- NULL # TODO: e.g. "RNA"
check_initial_seurat_save = FALSE # TODO: Set. If true, we will see if an initial Seurat object has been saved at `data/Processed_Datasets/{dataset_name}/`

# Load processing program
name_str <- paste0(dataset_name,"_",program_name)
program_filepath <- here::here("Code","Process_Datasets","Programs",paste0(name_str,".R"))
if(!file.exists(program_filepath)) { stop("No file provided for processing program") }
source(program_filepath)

# If a baseline Seurat object is created, it can be saved in a number of formats
# These include an `.Rds` file with the Seurat object, a `.Rds` file with a list form, an `.h5Seurat` object, and `SingleCellExperiment`
# The list and SCE formats are for use outside this pipeline and cannot be loaded in for reanalysis
# TODO: Pick formats to save initial Seuat object as
formats_to_save = list("seurat","list","h5Seurat","sce") # All options 

seurat_object <- NULL
initial_seurat_save_path <- here::here("data","Processed_Datasets",dataset_name,name_str,"initial_seurat")

initial_seurat_save_found <- FALSE # For printing in text
rds_save_used <- TRUE # For printing in text
# Rds appears to give faster load times than h5Seurat, so we check that format first.
if (check_initial_seurat_save && file.exists(paste0(initial_seurat_save_path,".Rds"))) {
	seurat_object <- readRDS(paste0(initial_seurat_save_path,".Rds"))
	initial_seurat_save_found <- TRUE
} else if (check_initial_seurat_save && file.exists(paste0(initial_seurat_save_path,".h5Seurat"))){
	seurat_object <- LoadH5Seurat(paste0(initial_seurat_save_path,".h5Seurat"))
	rds_save_used <- FALSE
	initial_seurat_save_found <- TRUE
} 

if (is.null(seurat_object)){ # Will still be NULL if saved data has not been loaded
	seurat_object <- build_seurat_object(dataset_name,program_name,primary_assay,save_formats=formats_to_save)
}

```

This notebook details the processing of the '`r dataset_name`' dataset through the '`r program_name`' processing program. The primary assay is `r primary_assay`.
The processing program which contains functions for generating a Seurat object from raw files and processing that object can be found at `Code/Process_Datasets/Programs/`r name_str`.R`.


`r if(!initial_seurat_save_found){"An unmodified Seurat object was computed de novo in this run."}`
`r if(!initial_seurat_save_found){paste0("It was saved in the following formats at Code/Process_Datasets/",dataset_name,"/",name_str,": ",toString(formats_to_save),".")}`

`r if(initial_seurat_save_found && rds_save_used){paste0("A previously computed and unmodified Seurat object was loaded from Code/Process_Datasets/",dataset_name,"/",name_str,"/initial_seurat.Rds.")}`
`r if(initial_seurat_save_found && !rds_save_used){paste0("A previously computed and unmodified h5Seurat object was loaded from Code/Process_Datasets/",dataset_name,"/",name_str,"/initial_seurat.h5Seurat.")}`



# Process Data

## Process Metadata

```{r process_metadata}
# There are two steps to process the metadata of the baseline Seurat object. Metadata is expected at the cellular level
# First, any custom metadata manipulation, such as combining two covariates into one, is performed
# Second, core covariates are mapped to a consistent name and variable class across all datasets
# The core covariates are: patient, group (phenotype), sample, cell_type, batch, num_genes, and num_reads
# It is the responsibility of the user to manually match names in the metadata data.frame to the standardized core covariate names

# TODO: Match covariate names from the dataset to the core covariates. Leave NULL if no corresponding column exists
patient_str <- NULL # The string of the patient ID covariate
group_str <- NULL # The group (phenotype/condition) covariate
sample_str <- NULL # The sample ID within a patient
cell_type_str <- NULL # The cell type covariate
batch_str <- NULL # The covariate associated with any possible batch effect such as processing site or date
num_features_str <- NULL # The covariate for the number of non-zero features in a given cell. `nFeature_RNA` is the Seurat default
num_reads_str <- NULL # The covariate for the number of reads in a given cell. `nCount_RNA` is the Seurat default

metadata_formatted <- format_metadata_df(seurat_object@meta.data,
 					 dataset=dataset_name,patient=patient_str, group=group_str,
 					 sample=sample_str,cell_type=cell_type_str,batch=batch_str,num_features=num_features_str,
 					 num_reads=num_reads_str,
 					 custom_steps=metadata_manipulation_fn)
```


```{r save_metadata,echo=FALSE}
# save the meta.data standalone for access without loading a Seurat object
saveRDS(metadata_formatted,file=here::here('data','Processed_Datasets',dataset_name,'metadata',paste0(name_str,"_metadata.Rds")))
seurat_object@meta.data <- metadata_formatted # update the pipeline Seurat object
```

```{r check_sample_bool,echo=FALSE}
single_sample_bool <- get_patient_sample_cross(seurat_object@meta.data)
```

`r paste0("There is ",ifelse(single_sample_bool,"","not "),"a single sample per patient.")`

The following function was used for dataset-specific pre-processing:

```{r print_meta_fn,echo=FALSE}
metadata_manipulation_fn
```

## Pre-Processing Pipeline

Before running the pre-processing pipeline there are `r dim(seurat_object)[2]` cells and `r dim(seurat_object)[1]` features in the dataset.

The following table shows the mapping from dataset specific covariates to their core covariate names. A missing variable name iditcates that covariate is not present.

| Core Covariate | Variable Name        |
|----------------|----------------------|
| Patient ID     | `r patient_str`      |
| Group          | `r group_str`        |
| Sample ID      | `r sample_str`       |
| Cell Type      | `r cell_type_str`    |
| Batch          | `r batch_str`        |
| Num. Features  | `r num_features_str` |
| Num. Reads     | `r num_reads_str`    |


The standard Seurat processing pipeline is normalization, variable feature selection, scaling, PCA, UMAP, and nearest neighbor estimation to enable clustering.

The following arguments are passed to each Seurat step, with NULL indicating an omitted step.

```{r print_pipeline,echo=FALSE,warning=FALSE}
seurat_params <- get_seurat_pipeline_params()
seurat_params
```

The custom functions dovetailed into the processing pipeline are as follows:

```{r dovetail_fns,echo=FALSE}
custom_fns <- get_seurat_pipeline_custom_fns()
custom_fns
```

```{r run_seurat,echo=FALSE}
#If the final object should be save in `h5Seurat` format, set the `save_h5` Boolean to True.
save_h5 <- FALSE

# run Seurat pipeline
pipeline_start <- proc.time()[3]
seurat_object <- run_seurat_pipeline(seurat_object,seurat_params,custom_fns)
pipeline_end <- proc.time()[3]
pipeline_delta <- pipeline_end - pipeline_start

if (save_h5) {SeuratDisk::SaveH5Seurat(seurat_object,filename=here::here("data","Processed_Datasets",dataset_name,name_str,paste0(name_str,".h5Seurat")),overwrite=TRUE)
} else { saveRDS(seurat_object, file=here::here("data","Processed_Datasets",dataset_name,name_str,paste0(name_str,".Rds")) ) }
```

After running the entire pipeline there are `r dim(seurat_object)[2]` cells and `r dim(seurat_object)[1]` features. It took `r pipeline_delta` seconds to run. 

The final Seurat object is saved at `r paste0("Code/Process_Datasets/",dataset_name,"/",name_str,"/",name_str,ifelse(save_h5,",h5Seurat",".Rds"))`


```{r get_props,echo=FALSE}
# For comparsion, we generate proprtion vectors for observtional units. A common choice is the proption of cell types in each sample
# TODO: Pick the aggregation and subset covariates
aggregation_variable <- NULL # e.g. "sample"
subset_variable <- NULL # e.g. "cell_type"

if (!is.null(aggregation_variable) && !is.null(subset_variable)){
	proportion_df <- get_cluster_props(seurat_object@meta.data,aggregation_variable,subset_variable) # The first covariate is the one to aggregate along, the second is the subsets (typically patients and cell types)
	saveRDS(proportion_df,file=here::here('data',"Processed_Datasets",dataset_name,'metadata',paste0(name_str,"_proportions.Rds")))
}
```

`r if (!is.null(aggregation_variable) && !is.null(subset_variable)) {paste0("For comparsion, we consider the proportion of ",subset_variable," in ",aggregation_variable," units. The first 5 units are shown.")}`
`r if (!is.null(aggregation_variable) && !is.null(subset_variable)) {paste0("The full tibble is saved at ",paste0("Code/Process_Datasets/",dataset_name,"/metadata/",paste0(name_str,"_proportions.Rds")))}`


```{r print_props,echo=FALSE}
if (!is.null(aggregation_variable) && !is.null(subset_variable)){
	print(proportion_df[1:5,])
}
```


# Exploratory Visualization

In this section we present a handful of basic visualizations of the results.

If cell type, group, and batch variables are present, we color our UMAP plots with those labels.

```{r umap_group,echo=FALSE}
plot_umap_group <- !all(is.na(seurat_object@meta.data$group))
if (plot_umap_group){
	plot <- Seurat::DimPlot(seurat_object,group.by="group") + labs(title = "UMAP Colored by Group Membership")
	plot
}
```


```{r umap_cell,echo=FALSE}
plot_umap_cell <- !all(is.na(seurat_object@meta.data$cell_type))
if (plot_umap_cell){
	plot <- Seurat::DimPlot(seurat_object,group.by="cell_type") + labs(title = "UMAP Colored by Cell Type")
	plot
}
```


```{r umap_batch,echo=FALSE}
plot_umap_batch <- !all(is.na(seurat_object@meta.data$batch))
if (plot_umap_batch){
	plot <- Seurat::DimPlot(seurat_object,group.by="batch") + labs(title = "UMAP Colored by Batch")
	plot
}
```

`r if (!is.null(aggregation_variable) && !is.null(subset_variable)){ paste0("The next histogram visualizes the cell type proportions for each unit. The units are ",aggregation_variable," and the subsets are ",subset_variable,".") }` 

```{r cluster_histogram,echo=FALSE}
if (!is.null(aggregation_variable) && !is.null(subset_variable)){
	proportion_long <- tibble::as_tibble(proportion_df,rownames="unit") %>% pivot_longer(cols=c(-unit),names_to="label",values_to="proportion")	
	prop_hist <- ggplot2::ggplot(proportion_long,aes(y=proportion,fill=label,x=unit)) + geom_bar(position="fill",stat="identity") +
		labs(title="Cluster Proportion Histogram",x=toString(aggregation_variable),fill=toString(subset_variable))+
		scale_x_discrete(guide = guide_axis(angle = -90))	
	print(prop_hist)
}
```

```{r get_custom_plots,echo=FALSE}
# print any custom plots as defined in the processing program
custom_plots <- get_custom_plots(seurat_object)
```

`r if(length(custom_plots)>0){ "The following plots are user specified and unique to this dataset" }`

```{r print_custom_plots,echo=FALSE}
if(length(custom_plots)>0){
	for (plot_index in 1:length(custom_plots)){
        	print(custom_plots[[plot_index]])
	}
}
```

# Documentation

Session Info:
```{r print_session_info}
print(utils::sessionInfo())
```

